---
title: "Prey Reduction Tech Memo"
author: "Meghan Roberts"
date: "2024-06-10"
output: html_document
---

This markdown contains analyses produced for the tech memo looking at cumulative prey reduction for SRKW across all fisheries by timestep. The figures are saved to saved_figures>tech_memo.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(janitor)
library(patchwork)
# library(ggtree)
library(stringr)
library(purrr)
library(png)
library(scatterpie)
library(grid)
library(treemapify)
library(cowplot)
```


```{r}
#read in data and align naming system
regions_sp<-read_sf(here("data","shapefiles", "Shelton_region_shapefile","Shelton_region_shapefile", "Shelton_regions_final.shp"))
regions_sp$layer[1]<-"Salish"
regions_sp$layer[2]<-"CA"
regions_sp$layer[3]<-"OR"
regions_sp$layer[5]<-"SWWCVI"
regions_sp<-regions_sp %>% rename("region"="layer")
data<-read_csv(here("data","USALLValid_TechMemo.csv"))
# write_csv(data, "tech_memo_data_nonshp.csv")

#join the data
data_spatial<-left_join(regions_sp, data)
# st_write(data_spatial, "tech_memo_data.shp")

```

#making a map
```{r}
#load in US state shapefile
US<-read_sf(here("data","shapefiles", "US","US", "States_shapefile.shp")) %>% 
  clean_names() %>% 
  filter(state_name!="ALASKA") %>% 
  filter(state_name!="HAWAII")

#load in Canada
CAN<-read_sf(here("data","shapefiles", "Canada","Canada", "gpr_000b11a_e.shp"))

#define colors
region_colors <- c("SWWCVI" = "#26456E", "Salish" = "#7AC7E2", "NOF" = "#64AAD2", 
                   "OR" = "#1C5A99", "CA" = "#B4D4DA")

map<-ggplot() +
  geom_sf(data=US, aes())+
  geom_sf(data=CAN, aes())+
  geom_sf(data = data_spatial, aes(fill=region)) +
  scale_fill_manual(values = region_colors) +  # Manual color assignment
  theme_minimal()+
  coord_sf(xlim = c(-129, -119), ylim = c(34, 51), expand = FALSE)+
  theme(legend.position = "none") + # Remove legend
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)
#two lines added to make blank map
        # ,legend.title=element_blank(),
        # legend.text = element_text(size = 14)
        )

#ggsave(here("saved_figures", "tech_memo", "region_map.jpg"), plot = map, height=10, width = 6)

map_legend<-ggplot() +
  geom_sf(data=US, aes())+
  geom_sf(data=CAN, aes())+
  geom_sf(data = data_spatial, aes(fill=region)) +
  scale_fill_manual(values = region_colors) +  # Manual color assignment
  theme_minimal()+
  coord_sf(xlim = c(-129, -119), ylim = c(34, 51), expand = FALSE)+
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

#make a corresponding bar plot
```{r}
bar<- ggplot(data_spatial, aes(x = timestep, y = avg_per_reduction, fill = region)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Timestep",
       y = "Average Percent Reduction",
       fill = "Region") +
  scale_fill_manual(values = region_colors) +
  theme_minimal()
```

#plot and map combo
```{r}
combined_plot <- map +bar
combined_plot

ggsave(here("saved_figures", "tech_memo", "regional_perc_reduction_tech_memo_1.jpg"), plot = combined_plot, width = 10, height = 5, dpi = 300)
```


#just 3 maps labeled with values
```{r}
#adding latitude and longitude for the labels
data_spatial$latitude=0
data_spatial$longitude=0
data_spatial$x_nudge=0
data_spatial$y_nudge=0
data_spatial <- data_spatial %>%
  mutate(latitude = ifelse(region == "CA", 125, latitude),
         longitude = ifelse(region == "CA", 35, longitude),
         x_nudge = ifelse(region == "CA", -1, x_nudge),
         y_nudge = ifelse(region == "CA", -4.5, y_nudge)) %>% 
  mutate(latitude = ifelse(region == "OR", 42.5, latitude),
         longitude = ifelse(region == "OR", 127, longitude),
         x_nudge = ifelse(region == "OR", -2.5, x_nudge),
         y_nudge = ifelse(region == "OR", -1, y_nudge)) %>% 
  mutate(latitude = ifelse(region == "NOF", 47, latitude),
         longitude = ifelse(region == "NOF", 127, longitude),
         x_nudge = ifelse(region == "NOF", -2.5, x_nudge),
         y_nudge = ifelse(region == "NOF", 0, y_nudge)) %>% 
  mutate(latitude = ifelse(region == "SWWCVI", 49, latitude),
         longitude = ifelse(region == "SWWCVI", 128, longitude),
         x_nudge = ifelse(region == "SWWCVI", -3, x_nudge),
         y_nudge = ifelse(region == "SWWCVI", 0.5, y_nudge)) %>% 
  mutate(latitude = ifelse(region == "Salish", 29, latitude),
         longitude = ifelse(region == "Salish", 122, longitude),
         x_nudge = ifelse(region == "Salish", 3, x_nudge),
         y_nudge = ifelse(region == "Salish", 0.4, y_nudge))
data_spatial$avg_per_reduction_label=NA

#breaking up data by timestep
oct_apr<-data_spatial %>% filter(timestep=="Oct-Apr")
may_jun<-data_spatial %>% filter(timestep=="May-Jun")
jul_sep<-data_spatial %>% filter(timestep=="Jul-Sep")

#getting the scale the same
# Generate gradient colors
gradient_colors <- scales::gradient_n_pal("Blues")(20)

# Define common scale limits based on the overall range of avg_per_reduction values
common_scale_limits <- range(c(oct_apr$avg_per_reduction, jul_sep$avg_per_reduction, may_jun$avg_per_reduction))
```

```{r}
oct_apr_map <- ggplot() +
  geom_sf(data = US, aes()) +
  geom_sf(data = CAN, aes()) +
  geom_sf(data = oct_apr, aes(fill = region)) +
  scale_fill_manual(values = region_colors) +
  theme_minimal() +
  coord_sf(xlim = c(-130, -119), ylim = c(34, 51), expand = FALSE) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  geom_sf_text(data = oct_apr, aes(label = avg_per_reduction_label),
               color = "black", size=3, fun.geometry = sf::st_centroid, 
               nudge_x=oct_apr$x_nudge, nudge_y = oct_apr$y_nudge)

oct_apr_map
```

```{r}
oct_apr_map <- ggplot() +
  geom_sf(data = US, aes()) +
  geom_sf(data = CAN, aes()) +
  geom_sf(data = oct_apr, aes(fill = avg_per_reduction)) +
  # scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", limits = common_scale_limits) +
  theme_minimal() +
  coord_sf(xlim = c(-133, -117), ylim = c(34, 51), expand = FALSE) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  geom_sf_text(data = oct_apr, aes(label = paste(region, " = \n", avg_per_reduction, "%")),
               color = "black", size=2.5, fun.geometry = sf::st_centroid, 
               nudge_x=oct_apr$x_nudge, nudge_y = oct_apr$y_nudge)+
  labs(x="", y="",fill="Average Percent Reduction", title="October - April")

# Print the map
# print(oct_apr_map)
```

```{r}
may_jun_map <- ggplot() +
  geom_sf(data = US, aes()) +
  geom_sf(data = CAN, aes()) +
  geom_sf(data = may_jun, aes(fill = avg_per_reduction)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", limits = common_scale_limits) +
  theme_minimal() +
  coord_sf(xlim = c(-133, -117), ylim = c(34, 51), expand = FALSE) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  geom_sf_text(data = may_jun, aes(label = paste(region, " = \n", avg_per_reduction, "%")),
               color = "black", size=2.5, fun.geometry = sf::st_centroid, 
               nudge_x=oct_apr$x_nudge, nudge_y = oct_apr$y_nudge)+
  labs(x="", y="",fill="Average Percent Reduction", title="May - June")

# Print the map
# print(may_jun_map)
```

```{r}
jul_sep_map <- ggplot() +
  geom_sf(data = US, aes()) +
  geom_sf(data = CAN, aes()) +
  geom_sf(data = jul_sep, aes(fill = avg_per_reduction)) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", limits = common_scale_limits) +
  theme_minimal() +
  coord_sf(xlim = c(-133, -117), ylim = c(34, 51), expand = FALSE) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  geom_sf_text(data = jul_sep, aes(label = paste(region, " = \n", avg_per_reduction, "%")),
               color = "black", size=2.5, fun.geometry = sf::st_centroid, 
               nudge_x=oct_apr$x_nudge, nudge_y = oct_apr$y_nudge)+
  labs(x="", y="",fill="Average Percent\nReduction", title="July - Sep")

# Print the map
# print(jul_sep_map)
```


#combine the maps
```{r}
perc_reduction_map<-oct_apr_map +may_jun_map +jul_sep_map
perc_reduction_map
# ggsave("regional_per_reduction_tech_memo.jpg", plot = perc_reduction_map, width = 10, height = 5, dpi = 300)
```

##Making a treemap instead of a pie chart
```{r}
#read in files from validation_run_data.Rmd

percent_reduction_all<-read_csv(here("data", "tech_memo","summary_data","Validation_all_regions_percent_reduction_1992-2020_tm.csv"))
percent_reduction_all_10yr<-read_csv(here("data", "tech_memo","summary_data","Validation_all_regions_percent_reduction_2011-2020_tm.csv"))

# Assuming percent_reduction_all has columns: time_step, avg_percent_reduction_Region, avg_percent_reduction_...

# Long format
percent_reduction_long <- percent_reduction_all %>%
  pivot_longer(cols = starts_with("avg_percent_reduction_"),
               names_to = "region",
               values_to = "percent_reduction") %>%
  clean_names()

percent_reduction_long$region <- str_replace_all(percent_reduction_long$region, "avg_percent_reduction_", "")

percent_reduction_long$scenario<-str_replace_all(percent_reduction_long$scenario, 
                                                 "Zero", "")

scenario_colors <- c("PS" = "#819171", "BC" = "#9CAF88", "PFMC" = "#DFE6DA", 
                     "SEAK" = "#CBD5B0")

percent_reduction_long<-percent_reduction_long %>% filter(scenario!="US")

#filter data
#CA
Cali_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="Cali")
Cali_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="Cali")
Cali_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="Cali")
#Salish
Salish_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="Salish")
Salish_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="Salish")
Salish_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="Salish")
#OR
OR_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="OR")
OR_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="OR")
OR_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="OR")
#NOF
NOF_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="NOF")
NOF_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="NOF")
NOF_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="NOF")
#SWWCVI
SWWCVI_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="SWWCVI")
SWWCVI_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="SWWCVI")
SWWCVI_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="SWWCVI")
```

#TS1 Treemaps for all years
```{r}
#with legend
#Cali
Cali_TS1_legend <- ggplot(Cali_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS1_tree <- ggplot(Cali_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS1_tree

OR_TS1_tree <- ggplot(OR_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS1_tree

NOF_TS1_tree <- ggplot(NOF_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS1_tree

Salish_TS1_tree <- ggplot(Salish_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS1_tree

SWWCVI_TS1_tree <- ggplot(SWWCVI_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS1_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS1_legend)
# Combine the plots using the + operator
TS1_combined <- Salish_TS1_tree+SWWCVI_TS1_tree+NOF_TS1_tree+OR_TS1_tree+Cali_TS1_tree+legend+
                  plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS1_combined <- TS1_combined +
  plot_annotation(title = "Average Percent Reduction in October to April by Region from 1992-2020")
                  
TS1_combined

ggsave(here("saved_figures", "tech_memo", "TS1_scenario_percent_reduction_1992_2020_tech_memo.jpg"), plot = TS1_combined, width = 8, height = 6, dpi = 300)

```


#TS2 Treemaps for all years 
```{r}
#with legend
#Cali
Cali_TS2_legend <- ggplot(Cali_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS2_tree <- ggplot(Cali_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS2_tree

OR_TS2_tree <- ggplot(OR_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS2_tree

NOF_TS2_tree <- ggplot(NOF_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS2_tree

Salish_TS2_tree <- ggplot(Salish_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS2_tree

SWWCVI_TS2_tree <- ggplot(SWWCVI_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS2_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS2_legend)
# Combine the plots using the + operator
TS2_combined <- Salish_TS2_tree+SWWCVI_TS2_tree+NOF_TS2_tree+OR_TS2_tree+Cali_TS2_tree+legend+
plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS2_combined <- TS2_combined +
  plot_annotation(title = "Average Percent Reduction in May to June by Region from 1992-2020")
                  
TS2_combined

ggsave(here("saved_figures", "tech_memo", "TS2_scenario_percent_reduction_1992_2020_tech_memo.jpg"), plot = TS2_combined, width = 8, height = 6, dpi = 300)

```


#TS3 Treemaps for all years 
```{r}
#with legend
#Cali
Cali_TS3_legend <- ggplot(Cali_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS3_tree <- ggplot(Cali_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS3_tree

OR_TS3_tree <- ggplot(OR_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS3_tree

NOF_TS3_tree <- ggplot(NOF_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS3_tree

Salish_TS3_tree <- ggplot(Salish_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS3_tree

SWWCVI_TS3_tree <- ggplot(SWWCVI_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS3_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS3_legend)
# Combine the plots using the + operator
TS3_combined <- Salish_TS3_tree+SWWCVI_TS3_tree+NOF_TS3_tree+OR_TS3_tree+Cali_TS3_tree+legend+
                  plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS3_combined <- TS3_combined +
  plot_annotation(title = "Average Percent Reduction in July to September by Region from 1992-2020")
TS3_combined

ggsave(here("saved_figures", "tech_memo", "TS3_scenario_percent_reduction_1992_2020_tech_memo.jpg"), plot = TS3_combined, width = 8, height = 6, dpi = 300)

```

##Repeating for the past 10 yrs and not all time

do not just save any of the files without running the whole script because things are named the same thing
```{r}
# Long format
percent_reduction_long <- percent_reduction_all_10yr %>%
  pivot_longer(cols = starts_with("avg_percent_reduction_"),
               names_to = "region",
               values_to = "percent_reduction") %>%
  clean_names()

percent_reduction_long$region <- str_replace_all(percent_reduction_long$region, "avg_percent_reduction_", "")

percent_reduction_long$scenario<-str_replace_all(percent_reduction_long$scenario, 
                                                 "Zero", "")

scenario_colors <- c("PS" = "#AA00D7", "BC" = "#DC86FA", "PFMC" = "#FBF2FF", 
                     "SEAK" = "#E2ABFF")

percent_reduction_long<-percent_reduction_long %>% filter(scenario!="US")

#filter data
#CA
Cali_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="Cali")
Cali_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="Cali")
Cali_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="Cali")
#Salish
Salish_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="Salish")
Salish_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="Salish")
Salish_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="Salish")
#OR
OR_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="OR")
OR_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="OR")
OR_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="OR")
#NOF
NOF_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="NOF")
NOF_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="NOF")
NOF_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="NOF")
#SWWCVI
SWWCVI_TS1<-percent_reduction_long %>% filter(time_step=="Oct_Apr") %>% 
  filter(region=="SWWCVI")
SWWCVI_TS2<-percent_reduction_long %>% filter(time_step=="May_Jun") %>% 
  filter(region=="SWWCVI")
SWWCVI_TS3<-percent_reduction_long %>% filter(time_step=="Jul_Sep") %>% 
  filter(region=="SWWCVI")
```


#TS1 Treemaps for 10yrs years
```{r}
#with legend
#Cali
Cali_TS1_legend <- ggplot(Cali_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS1_tree <- ggplot(Cali_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS1_tree

OR_TS1_tree <- ggplot(OR_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS1_tree

NOF_TS1_tree <- ggplot(NOF_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS1_tree

Salish_TS1_tree <- ggplot(Salish_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS1_tree

SWWCVI_TS1_tree <- ggplot(SWWCVI_TS1, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS1_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS1_legend)
# Combine the plots using the + operator
TS1_combined <- Salish_TS1_tree+SWWCVI_TS1_tree+NOF_TS1_tree+OR_TS1_tree+Cali_TS1_tree+legend+
                  plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS1_combined <- TS1_combined +
  plot_annotation(title = "Average Percent Reduction in October to April by Region from 2011-2020")
                  
TS1_combined

ggsave(here("saved_figures", "tech_memo", "TS1_scenario_percent_reduction_2011-2020_tech_memo.jpg"), plot = TS1_combined, width = 8, height = 6, dpi = 300)

```


#TS2 Treemaps for 10 years 
```{r}
#with legend
#Cali
Cali_TS2_legend <- ggplot(Cali_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS2_tree <- ggplot(Cali_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS2_tree

OR_TS2_tree <- ggplot(OR_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS2_tree

NOF_TS2_tree <- ggplot(NOF_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS2_tree

Salish_TS2_tree <- ggplot(Salish_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS2_tree

SWWCVI_TS2_tree <- ggplot(SWWCVI_TS2, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS2_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS2_legend)
# Combine the plots using the + operator
TS2_combined <- Salish_TS2_tree+SWWCVI_TS2_tree+NOF_TS2_tree+OR_TS2_tree+Cali_TS2_tree+legend+
plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS2_combined <- TS2_combined +
  plot_annotation(title = "Average Percent Reduction in May to June by Region from 2011-2020")
                  
TS2_combined

ggsave(here("saved_figures", "tech_memo", "TS2_scenario_percent_reduction_2011-2020_tech_memo.jpg"), plot = TS2_combined, width = 8, height = 6, dpi = 300)

```


#TS3 Treemaps for 10 years 
```{r}
#with legend
#Cali
Cali_TS3_legend <- ggplot(Cali_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "right") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS3_tree <- ggplot(Cali_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("CA")

Cali_TS3_tree

OR_TS3_tree <- ggplot(OR_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("OR")

OR_TS3_tree

NOF_TS3_tree <- ggplot(NOF_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("NOF")

NOF_TS3_tree

Salish_TS3_tree <- ggplot(Salish_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("Salish")

Salish_TS3_tree

SWWCVI_TS3_tree <- ggplot(SWWCVI_TS3, aes(area = percent_reduction, fill = scenario, label = paste0(percent_reduction, "%"))) +
  geom_treemap() +
  geom_treemap_text(colour = "black", size = 13) +
  scale_fill_manual(values = scenario_colors) +
  theme(legend.position = "none") +  # You can adjust the legend position as needed
  guides(fill = guide_legend(title = "Scenario"))+  # Set the legend title if needed
  ggtitle("SWWCVI")

SWWCVI_TS3_tree

#combine plots
grid_layout <- c(2, 3)

#get legend
legend <- cowplot::get_legend(Cali_TS3_legend)
# Combine the plots using the + operator
TS3_combined <- Salish_TS3_tree+SWWCVI_TS3_tree+NOF_TS3_tree+OR_TS3_tree+Cali_TS3_tree+legend+
                  plot_layout(ncol = grid_layout[2], byrow = TRUE)
TS3_combined <- TS3_combined +
  plot_annotation(title = "Average Percent Reduction in July to September by Region from 2011-2020")
TS3_combined

ggsave(here("saved_figures", "tech_memo", "TS3_scenario_percent_reduction_2011-2020_tech_memo.jpg"), plot = TS3_combined, width = 8, height = 6, dpi = 300)

```