---
title: "Validation Run Data"
author: "Meghan Roberts"
date: "2024-06-10"
output: html_document
---

This Rmarkdown contains analyses related to the validation run data provided by Jon Carey. The final chunk produces table_data which contains information used in the 2024 PSC 1-yr biop.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(ggplot2)
library(janitor)
library(tidyverse)
library(dplyr)
```

#Validation Output Analyses
####Loading in the data
```{r}
percent_reduction<-read_csv(here("data", "SRKW_output_Validation_1992-2020", "SRKW_output_Validation_1992-2020", "SRKW_reduction_percent_all_Valid2022.csv"))

starting_abundances<-read_csv(here("data", "SRKW_output_Validation_1992-2020", "SRKW_output_Validation_1992-2020", "SRKW_preFish_zero_all_Valid2022.csv"))

nominal_reduction<-read_csv(here("data", "SRKW_output_Validation_1992-2020", "SRKW_output_Validation_1992-2020", "SRKW_reduction_nominal_all_Valid2022.csv"))
```

###Tech Memo (all regions)
####Finding the average percent reduction by fishery 
```{r}
### getting the df in a workable format
#convert years to numeric
percent_reduction$Year<-as.numeric(percent_reduction$Year)
#remove percent sign
substitute_values <- function(column) {
  as.numeric(gsub("%", "", column))
}
# Apply the function to multiple columns
percent_reduction <- percent_reduction  %>%
  mutate_at(vars(Salish, OR, NOF, Cali, SWWCVI), list(~substitute_values(.)))

###group by and summarize
percent_reduction_all <- percent_reduction %>%
  group_by(TimeStep, Scenario) %>%
  summarise(
    avg_percent_reduction_Salish = round(mean(Salish, na.rm = TRUE),2),
    avg_percent_reduction_OR = round(mean(OR, na.rm = TRUE),2),
    avg_percent_reduction_NOF = round(mean(NOF, na.rm = TRUE),2),
    avg_percent_reduction_Cali = round(mean(Cali, na.rm = TRUE),2),
    avg_percent_reduction_SWWCVI = round(mean(SWWCVI, na.rm = TRUE),2),
    .groups = 'drop'
  )
write_csv(percent_reduction_all, here("data", "tech_memo","summary_data","Validation_all_regions_percent_reduction_1992-2020_tm.csv"))

percent_reduction_10yr<-percent_reduction %>% filter(Year>2010)

percent_reduction_all_10yr <- percent_reduction_10yr %>%
  group_by(TimeStep, Scenario) %>%
  summarise(
    avg_percent_reduction_Salish = round(mean(Salish, na.rm = TRUE),2),
    avg_percent_reduction_OR = round(mean(OR, na.rm = TRUE),2),
    avg_percent_reduction_NOF = round(mean(NOF, na.rm = TRUE),2),
    avg_percent_reduction_Cali = round(mean(Cali, na.rm = TRUE),2),
    avg_percent_reduction_SWWCVI = round(mean(SWWCVI, na.rm = TRUE),2),
    .groups = 'drop'
  )
write_csv(percent_reduction_all_10yr, here("data", "tech_memo","summary_data","Validation_all_regions_percent_reduction_2011-2020_tm.csv"))

## also create df from 1999-2020
percent_reduction_20yr<-percent_reduction %>% filter(Year>1998)

percent_reduction_all_10yr <- percent_reduction_10yr %>%
  group_by(TimeStep, Scenario) %>%
  summarise(
    avg_percent_reduction_Salish = round(mean(Salish, na.rm = TRUE),2),
    avg_percent_reduction_OR = round(mean(OR, na.rm = TRUE),2),
    avg_percent_reduction_NOF = round(mean(NOF, na.rm = TRUE),2),
    avg_percent_reduction_Cali = round(mean(Cali, na.rm = TRUE),2),
    avg_percent_reduction_SWWCVI = round(mean(SWWCVI, na.rm = TRUE),2),
    .groups = 'drop'
  )
write_csv(percent_reduction_all_10yr, here("data", "tech_memo","summary_data","Validation_all_regions_percent_reduction_2011-2020_tm.csv"))


## all years

percent_reduction_all_allyr <- percent_reduction %>%
  group_by(TimeStep, Scenario) %>%
  summarise(
    avg_percent_reduction_Salish = round(mean(Salish, na.rm = TRUE),2),
    avg_percent_reduction_OR = round(mean(OR, na.rm = TRUE),2),
    avg_percent_reduction_NOF = round(mean(NOF, na.rm = TRUE),2),
    avg_percent_reduction_Cali = round(mean(Cali, na.rm = TRUE),2),
    avg_percent_reduction_SWWCVI = round(mean(SWWCVI, na.rm = TRUE),2),
    .groups = 'drop'
  )
```

###1 yr Biop
####Finding the average percent reduction in Salish over all time and past 10 yrs by timestep
```{r}
### ALL YEARS

#filter
percent_reduction_salish_all<-percent_reduction %>% dplyr::select("Year", "Scenario", "TimeStep", "Salish")
#summarize results
percent_reduction_salish_all<-percent_reduction_salish_all %>% group_by(TimeStep) %>% 
  summarise(avg_percent_reduction = mean(Salish),.groups = 'drop')

### PAST 10 YRS

#filter to past 10yrs
percent_reduction_salish_10<-percent_reduction %>% dplyr::select("Year", "Scenario", "TimeStep", "Salish")%>% filter(Year>2010)
#summarize results
percent_reduction_salish_10<-percent_reduction_salish_10 %>% group_by(TimeStep) %>% 
  summarise(avg_per_reduction = mean(Salish),.groups = 'drop')

## PAST 20 YRS
percent_reduction_salish_20<-percent_reduction %>% dplyr::select("Year", "Scenario", "TimeStep", "Salish")%>% filter(Year>2000)
#summarize results
percent_reduction_salish_20<-percent_reduction_salish_20 %>% group_by(TimeStep, Scenario) %>% 
  summarise(avg_per_reduction = mean(Salish),.groups = 'drop')

```

####Finding the average nominal reduction in Salish over all time and past 10 yrs by timestep
```{r}
### ALL YEARS

#filter
nominal_reduction_salish_all<-nominal_reduction %>% dplyr::select("Year", "Scenario", "TimeStep", "Salish")
#summarize results
nominal_reduction_salish_all<-nominal_reduction_salish_all %>% group_by(TimeStep) %>% 
  summarise(avg_percent_reduction = mean(Salish),.groups = 'drop')

### PAST 10 YRS

#filter to past 10yrs
nominal_reduction_salish_10<-nominal_reduction %>% dplyr::select("Year", "Scenario", "TimeStep", "Salish")%>% filter(Year>2010)
#summarize results
nominal_reduction_salish_10<-nominal_reduction_salish_10 %>% group_by(TimeStep) %>% 
  summarise(avg_per_reduction = mean(Salish),.groups = 'drop')
```

####Finding the average starting abundance over all time and past 10 yrs
Only look at timestep 1, which is October - April
```{r}
starting_abundances_Salish<-starting_abundances %>% filter(TimeStep=="Oct_Apr") %>% 
  filter(Scenario=="ZeroSEAK") %>% 
  dplyr::select("Year", "Scenario", "Salish")

mean(starting_abundances_Salish$Salish)
starting_abundances_Salish_10yr<-starting_abundances_Salish %>% filter(Year>2010)
mean(starting_abundances_Salish_10yr$Salish)

ggplot(data=starting_abundances_Salish, aes(x=Year, y=Salish))+
  geom_point()+
  geom_line()+
  theme_minimal()+
  ylab("Number of Chinook in the Salish Sea")
```

####Table in Effects
```{r}
table_data<-starting_abundances_Salish %>% dplyr::select(-Scenario) %>% 
  rename("starting_abundance"="Salish")
nominal_reduction_table<-nominal_reduction %>% filter(Scenario=="ZeroPS") %>% 
  filter(TimeStep=="Oct_Apr") %>% dplyr::select(Year, Salish)%>% 
  rename("nominal_reduction"="Salish")
percent_reduction_table<-percent_reduction %>% filter(Scenario=="ZeroPS") %>% 
  filter(TimeStep=="Oct_Apr") %>% dplyr::select(Year, Salish)%>% 
  rename("percent_reduction"="Salish")

table_data<- table_data %>% right_join(nominal_reduction_table)
table_data<- table_data %>% right_join(percent_reduction_table)
```